\pagebreak
% Dedication
\clearpage
\vspace*{\fill}

{
\large
\hspace{3.5cm} \hfill \textit{Let this work stand not as a testament to the greatness of man, but to the greatness of God. Without His guidance and strength, this dissertation would not have taken the shape it holds today.\\[1em]
\noindent ``Let your light so shine before men, that they may see your good works and glorify your Father in heaven.''}
\textbf{-- Matthew 5:16}
}

\vspace*{\fill}

\clearpage

\chapter*{Acknowledgements}
% \addcontentsline{toc}{chapter}{Acknowledgements}
% \thispagestyle{empty}

\noindent While this journey is presented as a single document encompassing a year’s worth of work, to me it signifies something far greater. It marks a turning point--a period of profound personal transformation and growth in ways I had not anticipated. Though this dissertation bears the name of a single author, it was made possible through the invaluable support of many individuals. I dedicate these few pages to all those who enabled this transformation in me, and who supported me in discovering and nurturing my passion.

\bigskip

\noindent First and foremost, I would like to thank God. There were days when I felt alone, confused, and incapable of making further progress--when even writing a single line of code felt overwhelmingly difficult amid the mental and physical strains. Yet, during these moments, I experienced spiritual growth that gave me the courage to persevere. Even when the cognitive burden felt paralysing and clarity eluded me, I found inspiration and strength in the following verse:

\bigskip\bigskip

\noindent
\textit{``In the beginning was the Word, and the Word was with God, and the Word was God. He was with God in the beginning. Through him all things were made; without him nothing was made that has been made. In him was life, and that life was the light of all mankind. The light shines in the darkness, and the darkness has not overcome it.''} \textbf{-- John 1:1–3}

\bigskip\bigskip

\noindent Without His influence and steady guidance, this dissertation would not have taken the form it has today, nor would its direction and clarity have been fully realised.

\bigskip

\noindent Secondly, I wish to express my deepest gratitude to my supervisors, Dr. Dylan Seychell and Dr. Konstantinos Makantasis, for their guidance and support throughout this process. Beyond directing my work, they helped cultivate my understanding of scientific methodology and encouraged my growth as an independent researcher. I am sincerely thankful for the guidance they provided from their experience (privileged information), their thoughtful feedback, and their unwavering commitment through countless meetings and extensive correspondence.

\bigskip

\noindent I am also grateful to Mr. Gabriel Hili and Mr. Jonathan Attard for their technical support in configuring the hardware infrastructure provided by the Department of Artificial Intelligence at the University of Malta. Their assistance was instrumental in enabling the training of the models developed as part of this dissertation.

\bigskip

\noindent Finally, I extend heartfelt thanks to my family--my backbone, and the foundation of all I strive for. Their wisdom, encouragement, and constant support sustained me through the most difficult moments. To my mother, who has supported me every step of the way and ensured I had the resources I needed to continue my studies, and to my father, who consistently reminds me to remain humble--I am eternally grateful. To my grandparents, both living and departed, and to my grandaunts and granduncles: thank you for instilling in me enduring values and a sense of simplicity in service to others. To my uncles, aunts, and cousins: thank you for your love, encouragement, and for always being there.

\bigskip

\noindent Just as the following words from Scripture reminded me of God’s constant presence in times of trial, they also echo the steadfast love and support my family has shown me throughout this journey:

\bigskip\bigskip

\noindent
\textit{``But now, thus says the Lord, who created you, \ldots, and He who formed you, \ldots : ‘Fear not, for I have redeemed you; I have called you by your name; You are Mine. When you pass through the waters, I will be with you; And through the rivers, they shall not overflow you. When you walk through the fire, you shall not be burned, Nor shall the flame scorch you.’''} \textbf{-- Isaiah 43:1–2}

\pagebreak
% Dedication
\clearpage
\vspace*{\fill}

{
\large
\textit{``I was one way... and now I am completely different. And the thing that happened in between... was Him.''}
\textbf{-- Mary Magdalene, The Chosen}
}

\vspace*{\fill}

%------------------------------------------------------------------------------------------

% This is the Old one for Bachelors: We say our respect here: Thank you, Grazzi Sinjur Alla Ahfirli Sinjur Alla u Ismaghni Sinjur Alla. Grazzi hafna. Amen

% As I lay here before I have written a single word, I am filled with anxiousness upon this journey, and I turn to pray to God; God please give me the courage and hope necessary to __ this journey. And God answered my prayers by providing me with the necessary people to __

% \vspace*{\fill}
% \begin{center}
% \vspace*{0.05\textheight} % Adjust the vertical spacing as needed
% \Large \textit{In loving memory of those who could not witness this achievement:}
    
% % \vspace{0.1\textheight}    
    
% \vspace{0.03\textheight}
% \large
% \begin{tabular}{lc}
%     \textbf{Angela Ciantar} & 2024\\ \hline
%     \textbf{Pauline Cachia} & 2020\\ \hline
%     \textbf{Mary Rose Abela} & 2018\\ \hline
%     \textbf{Paul Bartolo} & 2017\\ \hline
%     \textbf{Josephine Bartolo} & 2014\\ \hline
%     \textbf{Maria Bartolo} & 2013\\ \hline
% \end{tabular}

% \vspace*{0.03\textheight} % Adjust the vertical spacing as needed
% \Large \textit{May their souls rest in peace.}

% \end{center}

% \vspace{0.4\textheight}

% \normalsize
% \noindent
% \textit{``Jesus looked at them and said, ``With man it is impossible, but not with God. For all things are possible with God.'''} \\[1ex] \noindent
% \rule{\textwidth}{1pt} \\[1ex]
% \hfill \textbf{-- Mark 10:27}



% \newpage

% While this journey is attributed to one author, it was made possible through the invaluable support of many others who contributed to its completion. I dedicate these few pages to all those who made a difference in this journey of discovery in my passion.

% \bigskip

% I extend my heartfelt dedication for this thesis primarily to my mother, Miriam Bartolo Abela, whose unwavering support and nurturing guidance have been instrumental in shaping my growth, providing constant encouragement, even during the most challenging moments when my resolve was tested to its limits.

% \bigskip

% I would also like to dedicate this thesis to my four grandparents: Catherine, and Victor Abela, Josephine, and Paul Bartolo, who were always there to support and instil upon me their good values and faith, reminding me of our family's humble beginnings.

% \bigskip

% I am deeply grateful to my overseers, supervisors, mentors, and those who provided invaluable opportunities for growth in the scientific field during the course of this thesis. Specifically, I extend my heartfelt thanks to my supervisors and mentors, Dr. Dylan Seychell, and Dr. Josef Bajada, who played integral roles in finishing this journey. I would also like to express my appreciation to Mr. Raymond Cuschieri, whose teachings and passion for his work sparked my love for programming and set me on this path.

% \bigskip

% I would also like to express my gratitude to my family for their unwavering support, providing me with shelter and comfort during difficult times. Specifically, I extend my thanks to Philip Abela, Antoine Abela, Nathalie Abela, Samuel Abela, Nazzareno Abela, Ivy Mae Dapar Banzon, Anthony Bartolo, Catherine Bartolo, Jimmy Bartolo, and Martina Bartolo.

% \bigskip
% \bigskip
% \bigskip
% \bigskip
% \bigskip
% \bigskip
% \bigskip
% \bigskip
% % \clearpage

% % \vspace{0.9\textheight}

% \noindent
% \textit{``Ask and it will be given to you; seek and you will find; knock and the door will be opened to you. For everyone who asks receives; the one who seeks finds; and to the one who knocks, the door will be opened.''}\\[1ex] \noindent
% \rule{\textwidth}{1pt} \\[1ex]
% \hfill \textbf{-- Matthew 7:7-8}

